---
title: "Hydra test comparisons"
format: html
editor: visual
---

## Build the data

Use `hydradata` package to create the .dat and .pin files. We are working from the forked `hydradata` repository at https://github.com/thefaylab/hydradata for this project. All of the input .csv files originally created in [this script](https://github.com/ices-eg/wg_WGSAM/blob/master/SkillAssessment/SkillAssessProject.Rmd) have been copied to the `data-raw` folder on thefaylab/hydradata repo. Note that some are still in use for variables ww are not changing. 

Now we are building inputs directly from the [`mskeyrun` data package](https://noaa-edab.github.io/ms-keyrun/) with a new function has been incorporated into the `hydradata` package [`hydradata::create_RData_mskeyrun(dattype = c("sim", "real"), nlenbin = 5)`](https://github.com/thefaylab/hydradata/blob/master/data-raw/create_RData_mskeyrun.R`).   

The new function allows us to specify whether we are building with simulated or real data, and the number of length bins to use in the input files. Default is 5 length bins but we will change this to generate different .dat and .pin files. 

The `create_RData_mskeyrun.R` file was sourced, and `create_RData_mskeyrun("sim", nlenbin = 5)` was run to make a new `hydraDataList_msk.rda` file in the package. Then the package is built locally, R is restarted, and the following code block is run to produce input files. 

We probably want to streamline the building process and keep separate hydradatalist R objects that have the different bin definitions. 

```{r, writedatfile5, eval=FALSE}

library(here)
library(hydradata)
inputs <- setup_default_inputs()
inputs$outDir <- here()

inputs$outputFilename <- "hydra_sim_NOBA_5bin" 

hydraDataList_5bin <- create_datpin_files(inputs,hydraDataList_msk)

# this saves the specific hydralist object, so we could saveRDS it to a diagnostics folder?
# advantage of rds format is we can assign it when reading it in to diagnostics scripts
saveRDS(hydraDataList_5bin, file.path(here("inputRdatalists/hydraDataList_5bin.rds")))

```

Run `create_RData_mskeyrun("sim", nlenbin = 10)` in `hydradata`, rebuild package, restart R, then...

```{r, writedatfile10, eval=FALSE}

library(here)
library(hydradata)
inputs <- setup_default_inputs()
inputs$outDir <- here()

inputs$outputFilename <- "hydra_sim_NOBA_10bin" 

hydraDataList_10bin <- create_datpin_files(inputs,hydraDataList_msk)

# this saves the specific hydralist object, so we could saveRDS it to a diagnostics folder?
# advantage of rds format is we can assign it when reading it in to diagnostics scripts
saveRDS(hydraDataList_10bin, file.path(here("inputRdatalists/hydraDataList_10bin.rds")))

```

Run `create_RData_mskeyrun("sim", nlenbin = 15)` in `hydradata`, rebuild package, restart R, then...

```{r, writedatfile15, eval=FALSE}

library(here)
library(hydradata)
inputs <- setup_default_inputs()
inputs$outDir <- here()

inputs$outputFilename <- "hydra_sim_NOBA_15bin" 

hydraDataList_15bin <- create_datpin_files(inputs,hydraDataList_msk)

# this saves the specific hydralist object, so we could saveRDS it to a diagnostics folder?
# advantage of rds format is we can assign it when reading it in to diagnostics scripts
saveRDS(hydraDataList_15bin, file.path(here("inputRdatalists/hydraDataList_15bin.rds")))

```

Run `create_RData_mskeyrun("sim", nlenbin = 20)` in `hydradata`, rebuild package, restart R, then...

```{r, writedatfile20, eval=FALSE}

library(here)
library(hydradata)
inputs <- setup_default_inputs()
inputs$outDir <- here()

inputs$outputFilename <- "hydra_sim_NOBA_20bin" 

hydraDataList_20bin <- create_datpin_files(inputs,hydraDataList_msk)

# this saves the specific hydralist object, so we could saveRDS it to a diagnostics folder?
# advantage of rds format is we can assign it when reading it in to diagnostics scripts
saveRDS(hydraDataList_20bin, file.path(here("inputRdatalists/hydraDataList_20bin.rds")))

```

Yes, slightly tedious, but doable.

Next up, real data. But first lets compare some simulated outputs changing the number of bins.

## Sarah's first run

Bring in code from [hydra_diag](https://github.com/thefaylab/hydra_diag):

```{r hydradiag}

# bring in reptoRlist
source("https://raw.githubusercontent.com/thefaylab/hydra_diag/main/R/read.report.R")

# read the rep file
test1 <- reptoRlist(here::here("hydra_sim.rep"))

names(test1)
```

Bring in inputs from the -ts.dat file (note that `hydradata` contains an R object with these inputs).

**ENSURE THE VERSION OF HYDRADATA MATCHES THE INPUTS** I saved the Rlist with the same name as dat and pin files so we can automate this

```{r readtsdat}

library(tidyverse)

#datfile <- here::here("hydra_sim_NOBA-ts.dat")

#write a function that splits the dat file appropriately
# e.g. read from Survey index to end survey index

#actually don't, use the list that built the dat file instead

# dattstoRlist <- function(fn, datn){
#   readLines(fn)  %>%          # Read file into a vector of strings
#   paste(collapse = "\n") %>%            # Stick them together with newlines
#   strsplit(paste0("# ", datn, ": +"))          # Split them into groups at the keyword
# 
# }
# 
# dat1 <- dattstoRlist(datfile)

datlist <- hydraDataList_5bin

obs_survey <- datlist$observedBiomass

obs_catch <- datlist$observedCatch

obs_diet <- datlist$observedSurvDiet




```

meantime copy paste portions of existing [diagnostics.R](https://github.com/thefaylab/hydra_diag/blob/main/Diagnostics/diagnostics.R)

but my report file is not the same?

```{r, eval=FALSE}
output <- test1

datfile <- here::here("hydra_sim_NOBA-ts.dat")

######### READ OBSERVED AND ESTIMATED SURVEY VALUES ###################

obs_survey<-read.table(datfile, skip=1695, nrows=1680, header=F)
colnames(obs_survey) <- c('number','year','species','type','InpN','1','2','3','4','5')
obs_survey<-obs_survey %>% pivot_longer(cols=6:10, names_to = "lenbin") %>%
  filter(value != -999)%>%         
  mutate(lenbin = as.integer(str_remove(lenbin, "V"))) 
obs_survey$label<-rep(("survey"),each=7189)
#dim(obs_survey) #7189

pred_surv<-output$pred_survey_size
obs_survey$pred_surv<-pred_surv
nll_survey<-output$nll_survey_size
obs_survey$nll_surv<-nll_survey
obs_survey$pearson<-((obs_survey$value-obs_survey$pred_surv)/sqrt(obs_survey$pred_surv))

colnames(obs_survey) <- c('number','year','species','type','InpN','lenbin','obs_value','label',
                          'pred_value','nll','pearson')


######### READ OBSERVED AND PREDICTED CATCH VALUES ###################

obs_catch<-read.table(datfile, skip=4109, nrows=608, header=F)
colnames(obs_catch) <- c('number','area','year','species','type','InpN','1','2','3','4','5')
obs_catch<-obs_catch %>% pivot_longer(cols=7:11, names_to = "lenbin") %>%
  mutate(lenbin = as.integer(str_remove(lenbin, "V"))) %>%
  filter(value != -999) 
obs_catch$label<-rep(("catch"),each=1855)
#dim(obs_catch) #1855

pred_catch<-output$pred_catch_size
obs_catch$pred_catch<-pred_catch
nll_catch<-output$nll_catch_size
obs_catch$nll_catch<-nll_catch
obs_catch$pearson<-((obs_catch$value-obs_catch$pred_catch)/sqrt(obs_catch$pred_catch))

colnames(obs_catch) <- c('number','area','year','species','type','InpN','lenbin','obs_value','label',
                         'pred_value','nll','pearson')


diet_catch<-full_join(obs_catch, obs_survey, by = NULL,copy = FALSE)

######### READ OBSERVED AND PREDICTED DIET PROPORTION VALUES ###################

obs_diet<-read.table(datfile, skip=4724, nrows=4721, header=F)
#obs_diet$label<-rep(("diet"),each=4721)
obs_diet<-obs_diet %>% pivot_longer(cols=6:17, names_to = "lenbin") %>%
  mutate(lenbin = as.integer(str_remove(lenbin, "V"))) %>%
  filter(value >0 ) 
obs_diet$label<-rep(("diet"),each=22810)
#dim(obs_diet) #22810

pred_diet<-output$pred_dietprop
obs_diet$pred_diet<-pred_diet
nll_diet<-output$nll_dietprop
obs_diet$nll_diet<-nll_diet
obs_diet$pearson<-((obs_diet$value-obs_diet$pred_diet)/sqrt(obs_diet$pred_diet))
colnames(obs_diet) <- c('number','year','species','lenbin','InpN','prey','obs_value','label', 'pred_value','nll','pearson')


mydata<-full_join(diet_catch, obs_diet, by = NULL,copy = FALSE)
mydata$prey[is.na(mydata$prey)] <- -99
mydata$area[is.na(mydata$area)] <- 1
mydata$type[is.na(mydata$type)] <- 0

```


## Try some plots

When you click the **Render** button a document will be generated that includes both content and the output of embedded code. You can embed code like this:

```{r}
1 + 1
```

You can add options to executable code like this

```{r}
#| echo: false
2 * 2
```

The `echo: false` option disables the printing of code (only output is displayed).
