---
title: "Hydra test comparisons"
format: html
editor: visual
---

## Build the data

Use `hydradata` package to create the .dat and .pin files. We are working from the forked `hydradata` repository at https://github.com/thefaylab/hydradata for this project. All of the input .csv files originally created in [this script](https://github.com/ices-eg/wg_WGSAM/blob/master/SkillAssessment/SkillAssessProject.Rmd) have been copied to the `data-raw` folder on thefaylab/hydradata repo. Note that some are still in use for variables ww are not changing.

Now we are building inputs directly from the [`mskeyrun` data package](https://noaa-edab.github.io/ms-keyrun/) with a new function has been incorporated into the `hydradata` package [`hydradata::create_RData_mskeyrun(dattype = c("sim", "real"), nlenbin = 5)`](https://github.com/thefaylab/hydradata/blob/master/data-raw/create_RData_mskeyrun.R%60).

The new function allows us to specify whether we are building with simulated or real data, and the number of length bins to use in the input files. Default is 5 length bins but we will change this to generate different .dat and .pin files.

The `create_RData_mskeyrun.R` file was sourced, and `create_RData_mskeyrun("sim", nlenbin = 5)` was run to make a new `hydraDataList_msk.rda` file in the package. Then the package is built locally, R is restarted, and the following code block is run to produce input files.

We probably want to streamline the building process and keep separate hydradatalist R objects that have the different bin definitions.

```{r, writedatfile5, eval=FALSE}

library(here)
library(hydradata)
inputs <- setup_default_inputs()
inputs$outDir <- here()

inputs$outputFilename <- "hydra_sim_NOBA_5bin_0comp" 

# tpl code removes 0 so replace in data--nope tpl now updated
# nbins <- hydraDataList_msk$Nsizebins
# hydraDataList_msk$observedCatchSize[,7:(6+nbins)][hydraDataList_msk$observedCatchSize[,7:(6+nbins)]==0] <- 1e-4
# hydraDataList_msk$observedSurvSize[,6:(5+nbins)][hydraDataList_msk$observedSurvSize[,6:(5+nbins)]==0] <- 1e-4

hydraDataList_5bin_0comp <- create_datpin_files(inputs,hydraDataList_msk)

# this saves the specific hydralist object, so we could saveRDS it to a diagnostics folder?
# advantage of rds format is we can assign it when reading it in to diagnostics scripts
saveRDS(hydraDataList_5bin_0comp, file.path(here("inputRdatalists/hydraDataList_5bin_0comp.rds")))

```

Run `create_RData_mskeyrun("sim", nlenbin = 10)` in `hydradata`, rebuild package, restart R, then...

```{r, writedatfile10, eval=FALSE}

library(here)
library(hydradata)
inputs <- setup_default_inputs()
inputs$outDir <- here()

inputs$outputFilename <- "hydra_sim_NOBA_10bin_0comp" 

# # tpl code removes 0 so replace in data
# nbins <- hydraDataList_msk$Nsizebins
# hydraDataList_msk$observedCatchSize[,7:(6+nbins)][hydraDataList_msk$observedCatchSize[,7:(6+nbins)]==0] <- 1e-4
# hydraDataList_msk$observedSurvSize[,6:(5+nbins)][hydraDataList_msk$observedSurvSize[,6:(5+nbins)]==0] <- 1e-4


hydraDataList_10bin_0comp <- create_datpin_files(inputs,hydraDataList_msk)

# this saves the specific hydralist object, so we could saveRDS it to a diagnostics folder?
# advantage of rds format is we can assign it when reading it in to diagnostics scripts
saveRDS(hydraDataList_10bin_0comp, file.path(here("inputRdatalists/hydraDataList_10bin_0comp.rds")))

```

Run `create_RData_mskeyrun("sim", nlenbin = 15)` in `hydradata`, rebuild package, restart R, then...

```{r, writedatfile15, eval=FALSE}

library(here)
library(hydradata)
inputs <- setup_default_inputs()
inputs$outDir <- here()

inputs$outputFilename <- "hydra_sim_NOBA_15bin" 

hydraDataList_15bin <- create_datpin_files(inputs,hydraDataList_msk)

# this saves the specific hydralist object, so we could saveRDS it to a diagnostics folder?
# advantage of rds format is we can assign it when reading it in to diagnostics scripts
saveRDS(hydraDataList_15bin, file.path(here("inputRdatalists/hydraDataList_15bin.rds")))

```

Run `create_RData_mskeyrun("sim", nlenbin = 20)` in `hydradata`, rebuild package, restart R, then...

```{r, writedatfile20, eval=FALSE}

library(here)
library(hydradata)
inputs <- setup_default_inputs()
inputs$outDir <- here()

inputs$outputFilename <- "hydra_sim_NOBA_20bin" 

hydraDataList_20bin <- create_datpin_files(inputs,hydraDataList_msk)

# this saves the specific hydralist object, so we could saveRDS it to a diagnostics folder?
# advantage of rds format is we can assign it when reading it in to diagnostics scripts
saveRDS(hydraDataList_20bin, file.path(here("inputRdatalists/hydraDataList_20bin.rds")))

```

Yes, slightly tedious, but doable.

Next up, real data. But first lets compare some simulated outputs changing the number of bins.

## 5 and 10 bins work, 15 and 20 dont

Model errors out. Terminal output:

    NECLNASARAHMAC:hydra_sim sarah.gaichas$ ./hydra_sim -ind hydra_sim_NOBA_15bin.dat -nohess -ainp hydra_sim_NOBA_15bin.pin
    q par map
    7
     1 7
     1 2 1
     1 3 1
     1 4 1
     1 6 1
     1 8 1
     1 10 1
     1 11 1
    1701
    completed predicted values for surveys
    starting commercial catch nll
    Error: Invalid index 25389 used for array range [1, 1] in "prevariable& dvar4_array::operator() (int i, int j, int k)".
    hslice index out of bounds

I wonder if the sizes are just too thin. Capelin is down to 2 cm size bins. Lets try 12 and see?

Run `create_RData_mskeyrun("sim", nlenbin = 12)` in `hydradata`, rebuild package, restart R, then...

```{r, writedatfile12, eval=FALSE}

library(here)
library(hydradata)
inputs <- setup_default_inputs()
inputs$outDir <- here()

inputs$outputFilename <- "hydra_sim_NOBA_12bin" 

hydraDataList_12bin <- create_datpin_files(inputs,hydraDataList_msk)

# this saves the specific hydralist object, so we could saveRDS it to a diagnostics folder?
# advantage of rds format is we can assign it when reading it in to diagnostics scripts
saveRDS(hydraDataList_12bin, file.path(here("inputRdatalists/hydraDataList_12bin.rds")))

```

Nope, same error. We will now use only 5 and 10 bins.

## Make real data take 1

Only 5 bins work at this time because we are using some of the original parameterization.

Run `create_RData_mskeyrun("real", nlenbin = 5)` in `hydradata`, rebuild package, restart R, then...

This now starts the time series in 1978 and replaces -999 with 0 in comps

```{r, writedatfilereal5, eval=FALSE}

library(here)
library(hydradata)
inputs <- setup_default_inputs()
inputs$outDir <- here()

inputs$outputFilename <- "GB-input/hydra_sim_GB_5bin_1978_inpN" 

hydraDataList_GB5bin_1978_inpN <- create_datpin_files(inputs,hydraDataList_msk)

# this saves the specific hydralist object, so we could saveRDS it to a diagnostics folder?
# advantage of rds format is we can assign it when reading it in to diagnostics scripts
saveRDS(hydraDataList_GB5bin_1978_inpN, file.path(here("inputRdatalists/hydraDataList_GB5bin_1978_inpN.rds")))

```


Make a new lower catch datafile for testing:

```{r, eval=FALSE}

library(here)
library(hydradata)
inputs <- setup_default_inputs()
inputs$outDir <- here()

inputs$outputFilename <- "hydra_sim_GB_5bin_lowcatch" 

hydraDataList_msk$observedCatch$catch <- hydraDataList_msk$observedCatch$catch/1000

hydraDataList_GB5bin_lowcatch <- create_datpin_files(inputs,hydraDataList_msk)

# this saves the specific hydralist object, so we could saveRDS it to a diagnostics folder?
# advantage of rds format is we can assign it when reading it in to diagnostics scripts
saveRDS(hydraDataList_GB5bin_lowcatch, file.path(here("inputRdatalists/hydraDataList_GB5bin_lowcatch.rds")))


```

OK. No catch and no predation

```{r, eval=FALSE}
library(here)
library(hydradata)
inputs <- setup_default_inputs()
inputs$outDir <- here()

inputs$outputFilename <- "hydra_sim_GB_5bin_nocatch_nopred" 

hydraDataList_msk$observedCatch$catch <- hydraDataList_msk$observedCatch$catch*0

hydraDataList_msk$foodweb <-  hydraDataList_msk$foodweb*0

hydraDataList_GB5bin_nocatch_nopred <- create_datpin_files(inputs,hydraDataList_msk)

# this saves the specific hydralist object, so we could saveRDS it to a diagnostics folder?
# advantage of rds format is we can assign it when reading it in to diagnostics scripts
saveRDS(hydraDataList_GB5bin_nocatch_nopred, file.path(here("inputRdatalists/hydraDataList_GB5bin_nocatch_nopred.rds")))

```


No catch, no predation, set Linf to max bin?

```{r, eval=FALSE}

library(here)
library(hydradata)
inputs <- setup_default_inputs()
inputs$outDir <- here()

inputs$outputFilename <- "hydra_sim_GB_5bin_nocatch_nopred" 

hydraDataList_msk$observedCatch$catch <- hydraDataList_msk$observedCatch$catch*0

hydraDataList_msk$foodweb <-  hydraDataList_msk$foodweb*0

hydraDataList_GB5bin_nocatch_nopred <- create_datpin_files(inputs,hydraDataList_msk)

# this saves the specific hydralist object, so we could saveRDS it to a diagnostics folder?
# advantage of rds format is we can assign it when reading it in to diagnostics scripts
saveRDS(hydraDataList_GB5bin_nocatch_nopred, file.path(here("inputRdatalists/hydraDataList_GB5bin_nocatch_nopred.rds")))


```

## New data inputs after WGSAM: 1 fleet per species

Also this dataset does not include length borrowing for fisheries

```{r, writedatfilereal5-Dec2022, eval=FALSE}

library(here)
library(hydradata)
inputs <- setup_default_inputs()
inputs$outDir <- here()

inputs$outputFilename <- "GB-input/hydra_sim_GB_5bin_1978_10F" 

hydraDataList_GB5bin_1978_inpN_10F <- create_datpin_files(inputs,hydraDataList_msk)

# this saves the specific hydralist object, so we could saveRDS it to a diagnostics folder?
# advantage of rds format is we can assign it when reading it in to diagnostics scripts
saveRDS(hydraDataList_GB5bin_1978_inpN_10F, file.path(here("inputRdatalists/hydraDataList_GB5bin_1978_10F.rds")))

```

## Can we use hydradata's plot functions?

Some work and create output in the results directory, needs more tweaking for current shape of -ts.dat inputs.

```{r, eval=FALSE}

hydradata::create_documentation(outFile = "documentation-5bin.html", outDir = here::here("results/05bin"), data = readRDS(here::here("inputRdatalists/hydraDataList_5bin.rds")))


```

## Sarah's first run, 5 length bins fully from mskeyrun source

Bring in code from [hydra_diag](https://github.com/thefaylab/hydra_diag):

```{r hydradiag}

# bring in reptoRlist
source("https://raw.githubusercontent.com/thefaylab/hydra_diag/main/R/read.report.R")

# read the rep file

output <- reptoRlist(here::here("results/05bin/hydra_sim.rep"))

names(output)
```

Bring in inputs from the -ts.dat file (note that `hydradata` contains an R object with these inputs). The `hydra_diag/hydra-diagnostics.rmd` now takes advantages of this.

**ENSURE THE VERSION OF HYDRADATA MATCHES THE INPUTS** I saved the Rlist with the same name as dat and pin files so we can automate this

```{r readtsdat}

library(tidyverse)

#hydraDataList <- readRDS(here::here("inputRdatalists/hydraDataList_GB5bin.rds"))
hydraDataList <- readRDS(here::here("inputRdatalists/hydraDataList_GB5bin_1978.rds"))
#repfile <- here::here("results/GB05bin/hydra_sim.rep")

obs_surveyB <- hydraDataList$observedBiomass %>%
  mutate(species = hydraDataList$speciesList[species],
         survey = as.factor(survey)) 


obs_catchB <- hydraDataList$observedCatch%>%
  mutate(species = hydraDataList$speciesList[species],
         fishery = as.factor(fishery))


#obs_diet <- datlist$observedSurvDiet #done below

```

Plot Survey B and Catch for the real data `inputRdatalists/hydraDataList_GB5bin.rds`

```{r plotdat}

ggplot(obs_surveyB, aes(x=year, y=biomass, colour = survey)) +
  geom_point() +
  #geom_errorbar() +
  theme_bw() +
  facet_wrap(~species, scales="free_y") +
  ggtitle("Survey input data by species")

ggplot(obs_catchB, aes(x=year, y=catch, colour = fishery)) +
  geom_point() +
  #geom_errorbar() +
  theme_bw() +
  facet_wrap(~species, scales="free_y") +
  ggtitle("Catch input data by species")


```

Function reptoRlist cannot read the tables of fits to survey and catch indices correctly.

Eventually fix in `hydra_sim` code report section, but for now read from already generated rep files:

```{r, eval=FALSE}

gettables <- function(repfile, biorows, catchrows){
  # scan repfile to the line "table of fits to survey"
  # next line is "survey biomass data, predicted, residual, nll"
  # columns are the -ts file inputs with predicted B, residual, and nll appended
  # based on https://stackoverflow.com/questions/37663246/extract-data-between-a-pattern-from-a-text-file
  reptext <- readLines(repfile)

  tablelist <- split(reptext, cumsum(grepl(pattern = 'table of fits to', reptext)))
  
  survlist <- as.list(tablelist$`1`[1:biorows])
  
  
  sv <- lapply(survlist[-c(1:2)], function(x) read.table(textConnection(x), header = FALSE))
  survdf <- do.call(rbind, sv)
  names(survdf) <- c("survey","year", "species", "biomass", "cv",
                     "pred_bio", "residual", "nll")

  catchlist <- as.list(tablelist$`2`[1:catchrows])
  
  ct <- lapply(catchlist[-c(1:2)], function(x) read.table(textConnection(x), header = FALSE))
  catchdf <- do.call(rbind, ct)
  names(catchdf) <- c("fishery", "area" ,   "year" ,   "species", "catch",  "cv",
                     "pred_catch", "residual", "nll")
  
  tablist <- list(survdf, catchdf)
  
  return(tablist)

}


biorows <- dim(obs_surveyB)[1]
catchrows <- dim(obs_catchB)[1]

indexfits <- gettables(repfile, biorows, catchrows)

```

## Plot indices

```{r, eval=FALSE}
#survey_obs <- hydraDataList$observedBiomass %>% 
survey_obspred <- indexfits[1][[1]] %>%
  mutate(obs = biomass + 1e-07,
         pred = pred_bio + 1e-07,
         log_obs = log(obs),
         log_pred = log(pred),
         log_lo = log_obs - 1.96*cv,
         log_hi = log_obs + 1.96*cv,
         obs_lo = exp(log_lo),
         obs_hi = exp(log_hi),
         species = hydraDataList$speciesList[species])
```

### Fits to survey indices of abundance {.tabset}

```{r, results = 'asis', eval=FALSE}
for (surv in unique(survey_obspred$survey)) {
  cat("#### Survey #", surv, " \n")
  p1 <- survey_obspred %>% 
  filter(survey == surv) %>% 
  ggplot() +
  aes(x= year, y = log_obs, group = species, col=factor(survey)) +
  geom_errorbar(aes(ymin = log_lo, ymax = log_hi)) +
  geom_point() +
  geom_line(aes(x=year, y=log_pred), col = "blue") +
  facet_wrap(~species, scales = "free_y") +
  theme_bw() +
  guides(species = "None")
  print(p1)
  cat(" \n\n")
  }
```

###  {.unnumbered}

```{r, eval=FALSE}
#catch_obs <- hydraDataList$observedCatch %>% 
catch_obspred <- indexfits[2][[1]] %>% 
  mutate(obs = catch + 1e-07,
         pred = pred_catch + 1e-07,
         log_obs = log(obs),
         log_pred = log(pred),
         log_lo = log_obs - 1.96*cv,
         log_hi = log_obs + 1.96*cv,
         obs_lo = exp(log_lo),
         obs_hi = exp(log_hi),
         species = hydraDataList$speciesList[species])
```

### fits to Catch time series {.tabset}

```{r, results = 'asis', eval=FALSE}
for (fleet in unique(catch_obspred$fishery)) {
  cat("#### Fleet #", fleet, " \n")
p3 <- catch_obspred %>% 
  filter(fishery == fleet,
         catch > 0) %>% 
  ggplot() +
  aes(x= year, y = log_obs, group = species, col= factor(fishery)) +
  geom_errorbar(aes(ymin = log_lo, ymax = log_hi)) +
  geom_point() +
  geom_line(aes(x=year, y=log_pred), col = "blue") +
  facet_wrap(~species, scales = "free_y") +
  theme_bw() +
  guides(species = "None")
  print(p3)
  cat(" \n\n")
}
```

###  {.unnumbered}

Above does indices, below does comps

Actually did these in the `hydra-diag` repo. Dont reproduce here.

Make this a function in `hydra_diag/hydra-diagnostics.rmd` currently at lines 122-211

