---
title: "Hydra test comparisons"
format: html
editor: visual
---

## Sarah's first run

Bring in code from [hydra_diag](https://github.com/thefaylab/hydra_diag):

```{r hydradiag}

# bring in reptoRlist
source("https://raw.githubusercontent.com/thefaylab/hydra_diag/main/Diagnostics/read.report.R")

# read the rep file
test1 <- reptoRlist("hydra_sim.rep")

names(test1)
```

Bring in inputs from the -ts.dat file:

```{r readtsdat}

library(tidyverse)

#write a function that splits the dat file appropriately
# e.g. read from Survey index to end survey index

dattstoRlist <- function(fn, datn){
  readLines(fn)  %>%          # Read file into a vector of strings
  paste(collapse = "\n") %>%            # Stick them together with newlines
  strsplit(paste0("# ", datn, ": +"))          # Split them into groups at the keyword

}



dat1 <- dattstoRlist("hydra_sim_NOBA-ts.dat")

```

## Try some plots

When you click the **Render** button a document will be generated that includes both content and the output of embedded code. You can embed code like this:

```{r}
1 + 1
```

You can add options to executable code like this

```{r}
#| echo: false
2 * 2
```

The `echo: false` option disables the printing of code (only output is displayed).
