---
title: "Hydra test comparisons"
format: html
editor: visual
---

## Build the data

Use `hydradata` package to create the .dat and .pin files. We are working from the forked `hydradata` repository at https://github.com/thefaylab/hydradata for this project. All of the input .csv files originally created in [this script](https://github.com/ices-eg/wg_WGSAM/blob/master/SkillAssessment/SkillAssessProject.Rmd) have been copied to the `data-raw` folder on thefaylab/hydradata repo. Note that some are still in use for variables ww are not changing.

Now we are building inputs directly from the [`mskeyrun` data package](https://noaa-edab.github.io/ms-keyrun/) with a new function has been incorporated into the `hydradata` package [`hydradata::create_RData_mskeyrun(dattype = c("sim", "real"), nlenbin = 5)`](https://github.com/thefaylab/hydradata/blob/master/data-raw/create_RData_mskeyrun.R%60).

The new function allows us to specify whether we are building with simulated or real data, and the number of length bins to use in the input files. Default is 5 length bins but we will change this to generate different .dat and .pin files.

The `create_RData_mskeyrun.R` file was sourced, and `create_RData_mskeyrun("sim", nlenbin = 5)` was run to make a new `hydraDataList_msk.rda` file in the package. Then the package is built locally, R is restarted, and the following code block is run to produce input files.

We probably want to streamline the building process and keep separate hydradatalist R objects that have the different bin definitions.

```{r, writedatfile5, eval=FALSE}

library(here)
library(hydradata)
inputs <- setup_default_inputs()
inputs$outDir <- here()

inputs$outputFilename <- "hydra_sim_NOBA_5bin" 

hydraDataList_5bin <- create_datpin_files(inputs,hydraDataList_msk)

# this saves the specific hydralist object, so we could saveRDS it to a diagnostics folder?
# advantage of rds format is we can assign it when reading it in to diagnostics scripts
saveRDS(hydraDataList_5bin, file.path(here("inputRdatalists/hydraDataList_5bin.rds")))

```

Run `create_RData_mskeyrun("sim", nlenbin = 10)` in `hydradata`, rebuild package, restart R, then...

```{r, writedatfile10, eval=FALSE}

library(here)
library(hydradata)
inputs <- setup_default_inputs()
inputs$outDir <- here()

inputs$outputFilename <- "hydra_sim_NOBA_10bin" 

hydraDataList_10bin <- create_datpin_files(inputs,hydraDataList_msk)

# this saves the specific hydralist object, so we could saveRDS it to a diagnostics folder?
# advantage of rds format is we can assign it when reading it in to diagnostics scripts
saveRDS(hydraDataList_10bin, file.path(here("inputRdatalists/hydraDataList_10bin.rds")))

```

Run `create_RData_mskeyrun("sim", nlenbin = 15)` in `hydradata`, rebuild package, restart R, then...

```{r, writedatfile15, eval=FALSE}

library(here)
library(hydradata)
inputs <- setup_default_inputs()
inputs$outDir <- here()

inputs$outputFilename <- "hydra_sim_NOBA_15bin" 

hydraDataList_15bin <- create_datpin_files(inputs,hydraDataList_msk)

# this saves the specific hydralist object, so we could saveRDS it to a diagnostics folder?
# advantage of rds format is we can assign it when reading it in to diagnostics scripts
saveRDS(hydraDataList_15bin, file.path(here("inputRdatalists/hydraDataList_15bin.rds")))

```

Run `create_RData_mskeyrun("sim", nlenbin = 20)` in `hydradata`, rebuild package, restart R, then...

```{r, writedatfile20, eval=FALSE}

library(here)
library(hydradata)
inputs <- setup_default_inputs()
inputs$outDir <- here()

inputs$outputFilename <- "hydra_sim_NOBA_20bin" 

hydraDataList_20bin <- create_datpin_files(inputs,hydraDataList_msk)

# this saves the specific hydralist object, so we could saveRDS it to a diagnostics folder?
# advantage of rds format is we can assign it when reading it in to diagnostics scripts
saveRDS(hydraDataList_20bin, file.path(here("inputRdatalists/hydraDataList_20bin.rds")))

```

Yes, slightly tedious, but doable.

Next up, real data. But first lets compare some simulated outputs changing the number of bins.

## 5 and 10 bins work, 15 and 20 dont

Model errors out. Terminal output:

    NECLNASARAHMAC:hydra_sim sarah.gaichas$ ./hydra_sim -ind hydra_sim_NOBA_15bin.dat -nohess -ainp hydra_sim_NOBA_15bin.pin
    q par map
    7
     1 7
     1 2 1
     1 3 1
     1 4 1
     1 6 1
     1 8 1
     1 10 1
     1 11 1
    1701
    completed predicted values for surveys
    starting commercial catch nll
    Error: Invalid index 25389 used for array range [1, 1] in "prevariable& dvar4_array::operator() (int i, int j, int k)".
    hslice index out of bounds

I wonder if the sizes are just too thin. Capelin is down to 2 cm size bins. Lets try 12 and see?

Run `create_RData_mskeyrun("sim", nlenbin = 12)` in `hydradata`, rebuild package, restart R, then...

```{r, writedatfile12, eval=FALSE}

library(here)
library(hydradata)
inputs <- setup_default_inputs()
inputs$outDir <- here()

inputs$outputFilename <- "hydra_sim_NOBA_12bin" 

hydraDataList_12bin <- create_datpin_files(inputs,hydraDataList_msk)

# this saves the specific hydralist object, so we could saveRDS it to a diagnostics folder?
# advantage of rds format is we can assign it when reading it in to diagnostics scripts
saveRDS(hydraDataList_12bin, file.path(here("inputRdatalists/hydraDataList_12bin.rds")))

```

Nope, same error. We will now use only 5 and 10 bins.

## Make real data take 1

Only 5 bins work at this time because we are using some of the original parameterization. 

Run `create_RData_mskeyrun("real", nlenbin = 5)` in `hydradata`, rebuild package, restart R, then...

```{r, writedatfilereal5, eval=FALSE}

library(here)
library(hydradata)
inputs <- setup_default_inputs()
inputs$outDir <- here()

inputs$outputFilename <- "hydra_sim_GB_5bin" 

hydraDataList_GB5bin <- create_datpin_files(inputs,hydraDataList_msk)

# this saves the specific hydralist object, so we could saveRDS it to a diagnostics folder?
# advantage of rds format is we can assign it when reading it in to diagnostics scripts
saveRDS(hydraDataList_GB5bin, file.path(here("inputRdatalists/hydraDataList_GB5bin.rds")))

```

## Can we use hydradata's plot functions?

Some work and create output in the results directory, needs more tweaking for current shape of -ts.dat inputs.

```{r, eval=FALSE}

hydradata::create_documentation(outFile = "documentation-5bin.html", outDir = here::here("results/05bin"), data = readRDS(here::here("inputRdatalists/hydraDataList_5bin.rds")))


```

## Sarah's first run, 5 length bins fully from mskeyrun source

Bring in code from [hydra_diag](https://github.com/thefaylab/hydra_diag):

```{r hydradiag}

# bring in reptoRlist
source("https://raw.githubusercontent.com/thefaylab/hydra_diag/main/R/read.report.R")

# read the rep file

output <- reptoRlist(here::here("results/05bin/hydra_sim.rep"))

names(output)
```

Bring in inputs from the -ts.dat file (note that `hydradata` contains an R object with these inputs). The `hydra_diag/hydra-diagnostics.rmd` now takes advantages of this.

**ENSURE THE VERSION OF HYDRADATA MATCHES THE INPUTS** I saved the Rlist with the same name as dat and pin files so we can automate this

```{r readtsdat}

library(tidyverse)

hydraDataList <- readRDS(here::here("inputRdatalists/hydraDataList_5bin.rds"))
repfile <- here::here("results/05bin/hydra_sim.rep")

obs_surveyB <- hydraDataList$observedBiomass

obs_catchB <- hydraDataList$observedCatch

#obs_diet <- datlist$observedSurvDiet #done below

```

Function reptoRlist cannot read the tables of fits to survey and catch indices correctly.

Eventually fix in `hydra_sim` code report section, but for now read from already generated rep files:

```{r}

gettables <- function(repfile, biorows, catchrows){
  # scan repfile to the line "table of fits to survey"
  # next line is "survey biomass data, predicted, residual, nll"
  # columns are the -ts file inputs with predicted B, residual, and nll appended
  # based on https://stackoverflow.com/questions/37663246/extract-data-between-a-pattern-from-a-text-file
  reptext <- readLines(repfile)

  tablelist <- split(reptext, cumsum(grepl(pattern = 'table of fits to', reptext)))
  
  survlist <- as.list(tablelist$`1`[1:biorows])
  
  
  sv <- lapply(survlist[-c(1:2)], function(x) read.table(textConnection(x), header = FALSE))
  survdf <- do.call(rbind, sv)
  names(survdf) <- c("survey","year", "species", "biomass", "cv",
                     "pred_bio", "residual", "nll")

  catchlist <- as.list(tablelist$`2`[1:catchrows])
  
  ct <- lapply(catchlist[-c(1:2)], function(x) read.table(textConnection(x), header = FALSE))
  catchdf <- do.call(rbind, ct)
  names(catchdf) <- c("fishery", "area" ,   "year" ,   "species", "catch",  "cv",
                     "pred_catch", "residual", "nll")
  
  tablist <- list(survdf, catchdf)
  
  return(tablist)

}


biorows <- dim(obs_surveyB)[1]
catchrows <- dim(obs_catchB)[1]

indexfits <- gettables(repfile, biorows, catchrows)

```

## Plot indices

```{r}
#survey_obs <- hydraDataList$observedBiomass %>% 
survey_obspred <- indexfits[1][[1]] %>%
  mutate(obs = biomass + 1e-07,
         pred = pred_bio + 1e-07,
         log_obs = log(obs),
         log_pred = log(pred),
         log_lo = log_obs - 1.96*cv,
         log_hi = log_obs + 1.96*cv,
         obs_lo = exp(log_lo),
         obs_hi = exp(log_hi),
         species = hydraDataList$speciesList[species])
```

### Fits to survey indices of abundance {.tabset}

```{r, results = 'asis'}
for (surv in unique(survey_obspred$survey)) {
  cat("#### Survey #", surv, " \n")
  p1 <- survey_obspred %>% 
  filter(survey == surv) %>% 
  ggplot() +
  aes(x= year, y = log_obs, group = species, col=factor(survey)) +
  geom_errorbar(aes(ymin = log_lo, ymax = log_hi)) +
  geom_point() +
  geom_line(aes(x=year, y=log_pred), col = "blue") +
  facet_wrap(~species, scales = "free_y") +
  theme_bw() +
  guides(species = "None")
  print(p1)
  cat(" \n\n")
  }
```

###  {.unnumbered}

```{r}
#catch_obs <- hydraDataList$observedCatch %>% 
catch_obspred <- indexfits[2][[1]] %>% 
  mutate(obs = catch + 1e-07,
         pred = pred_catch + 1e-07,
         log_obs = log(obs),
         log_pred = log(pred),
         log_lo = log_obs - 1.96*cv,
         log_hi = log_obs + 1.96*cv,
         obs_lo = exp(log_lo),
         obs_hi = exp(log_hi),
         species = hydraDataList$speciesList[species])
```

### fits to Catch time series {.tabset}

```{r, results = 'asis'}
for (fleet in unique(catch_obspred$fishery)) {
  cat("#### Fleet #", fleet, " \n")
p3 <- catch_obspred %>% 
  filter(fishery == fleet,
         catch > 0) %>% 
  ggplot() +
  aes(x= year, y = log_obs, group = species, col= factor(fishery)) +
  geom_errorbar(aes(ymin = log_lo, ymax = log_hi)) +
  geom_point() +
  geom_line(aes(x=year, y=log_pred), col = "blue") +
  facet_wrap(~species, scales = "free_y") +
  theme_bw() +
  guides(species = "None")
  print(p3)
  cat(" \n\n")
}
```

###  {.unnumbered}

Above does indices, below does comps

Actually did these in the `hydra-diag` repo. Dont reproduce here.

Make this a function in `hydra_diag/hydra-diagnostics.rmd` currently at lines 122-211

```{r joinobspred}

obs_survey <- hydraDataList$observedSurvSize %>% tibble()
#colnames(obs_survey) <- c('number','year','species','type','InpN','1','2','3','4','5')  #todo - modify so number of bins is based on size of dataframe
obs_survey <- obs_survey %>% pivot_longer(cols=6:ncol(.), names_to = "lenbin") %>%
  filter(value != -999)%>%
  mutate(lenbin = as.integer(str_remove(lenbin, "sizebin")),
         label = rep("survey",nrow(.)),
         species = hydraDataList$speciesList[species])
#obs_survey$label<-rep(("survey"),each=7189)
#dim(obs_survey) #7189

pred_surv<-output$pred_survey_size
obs_survey$pred_surv<-pred_surv
nll_survey<-output$nll_survey_size
obs_survey$nll_surv<-nll_survey
obs_survey$pearson<-((obs_survey$value-obs_survey$pred_surv)/sqrt(obs_survey$pred_surv))

colnames(obs_survey) <- c('number','year','species','type','InpN','lenbin','obs_value','label',
                          'pred_value','nll','pearson')

#obs_catch<-read.table(tsfile, skip=4109, nrows=608, header=F)
obs_catch <- hydraDataList$observedCatchSize %>% tibble()
#colnames(obs_catch) <- c('number','area','year','species','type','InpN','1','2','3','4','5')
obs_catch<-obs_catch %>% pivot_longer(cols=7:ncol(.), names_to = "lenbin") %>%
  mutate(lenbin = as.integer(str_remove(lenbin, "sizebin")),
         label = rep("catch",nrow(.)),
         species = hydraDataList$speciesList[species]) %>%
  filter(value != -999)
#obs_catch$label<-rep(("catch"),each=1855)
#dim(obs_catch) #1855

pred_catch<-output$pred_catch_size
obs_catch$pred_catch<-pred_catch
nll_catch<-output$nll_catch_size
obs_catch$nll_catch<-nll_catch
obs_catch$pearson<-((obs_catch$value-obs_catch$pred_catch)/sqrt(obs_catch$pred_catch))

colnames(obs_catch) <- c('number','area','year','species','type','InpN','lenbin','obs_value','label',
                         'pred_value','nll','pearson')


#diet_catch<-full_join(obs_catch, obs_survey, by = NULL,copy = FALSE)
diet_catch <- bind_rows(obs_catch, obs_survey)

#### READ OBSERVED (.dat) AND PREDICTED (.rep) DIET PROPORTION VALUES ####
# todo as above, make the data file an argument and have the dimensions be determined from the file rather than hard-coded
#obs_diet<-read.table(tsfile, skip=4724, nrows=4721, header=F)
obs_diet <- hydraDataList$observedSurvDiet %>% tibble()

#obs_diet$label<-rep(("diet"),each=4721)
obs_diet<-obs_diet %>% pivot_longer(cols=6:ncol(.), names_to = "prey") %>%
  mutate(#lenbin = as.integer(str_remove(lenbin, "V")),
         species = hydraDataList$speciesList[species],
         label = rep("diet",nrow(.))) %>%
  filter(value >0 )
#obs_diet$label<-rep(("diet"),each=22810)
#dim(obs_diet) #22810

pred_diet<-output$pred_dietprop
obs_diet$pred_diet<-pred_diet
nll_diet<-output$nll_dietprop
obs_diet$nll_diet<-nll_diet
obs_diet$pearson<-((obs_diet$value-obs_diet$pred_diet)/sqrt(obs_diet$pred_diet))
colnames(obs_diet) <- c('number','year','species','lenbin','InpN','prey','obs_value','label', 'pred_value','nll','pearson')

#mydata<-full_join(diet_catch, obs_diet, by = NULL,copy = FALSE)
#mydata <- bind_rows(diet_catch, obs_diet)
mydata <- diet_catch %>%
  mutate(#prey = replace_na(prey, -99),
         area = replace_na(area, 1),
         type = replace_na(type, 0))
data <- mydata
data$residual<-ifelse(data$pearson<0,"negative","positive")
data$res_abs<-abs(data$pearson)
data<-as.data.frame(data)
#head(data)

#names(data)
#str(data)

#select type of data "label= catch, survey or diet"
temp.catch = data[which(data$label == "catch"),]
temp.surv = data[which(data$label == "survey"),]
#temp.diet = data[which(data$label == "diet"),]

data <- obs_diet #data <- read.csv("outputs/survdiet.csv", header = T)
#data<- select(data, -X)
data$residual<-ifelse(data$pearson<0,"negative","positive")
data$res_abs<-abs(data$pearson)
temp.diet = data[which(data$label == "diet"),]


```
